---
title: "Stat Inference Project Part 1"
author: "Joe Logan"
output:
  html_document: default
  pdf_document: default
---

## Statistical Inference Final Project 
### Part 1: Simulation Exercise 

In this project I will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution will be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda. I will set lambda = 0.2 for all of the simulations. Finally, I will investigate the distribution of averages of 40 exponentials. Note that I will need to do a thousand simulations.

 ********  

### Initialization of required libraries and set Options
```{r include=FALSE, cache=FALSE}
require(knitr)
require(ggplot2)
knitr::opts_chunk$set(echo = TRUE)
```


### Setup the simulation
```{r cache=FALSE, message=FALSE}
# Set the constants given in the problem statement
lambda <- 0.2 # given
nExp <- 40 # number of exponentials (given)
nSimulations <- 1000 # number of exponentials (given)

set.seed(567)

# run the test resulting in n x numberOfSimulations matrix
exponentialDistributions <- matrix(data=rexp(nExp * nSimulations, lambda), nrow=nSimulations)
exponentialDistributionMeans <- data.frame(means=apply(exponentialDistributions, 1, mean))
actualMean <- mean(exponentialDistributionMeans$means)

```

### Plot the data
```{r cache=FALSE, message=FALSE}
# plot the means
ggplot(data = exponentialDistributionMeans, aes(x = means)) + 
        geom_histogram(binwidth=lambda, fill="blue",color="black") +   
        labs(title="Distribution of the mean of 40 Exponentials", x="Mean of 40 Selections", y="Density") +
        geom_vline(xintercept=actualMean,size=1.0, color="red") # add a line for the actual mean
        scale_x_continuous(breaks=round(seq(min(exponentialDistributionMeans$means), max(exponentialDistributionMeans$means), by=1)))
```

*********

### Comparing the theoretical mean to the sample (actual) mean
#### By definition, theoretical mean is 1/lambda
```{r cache=FALSE, message=FALSE}
# Calculating the theoretical mean
mu <- 1/lambda
```

#### Calculating the actual mean
```{r cache=FALSE, message=FALSE}
actualMean <- mean(exponentialDistributionMeans$means)
```

The theoretical mean is `r mu ` and the sample mean is `r actualMean`  
We can see that these values are very close.

### Comparing the theoretical variance to the sample variance
#### R is equiped with the var function which can give us our sample variance
```{r cache=FALSE, message=FALSE}
# Compute the sample variance
sampleVariance <- var(as.vector(exponentialDistributionMeans$means))

# Give the theoretical variance
# We get this by dividing the mean by the square root of our number of exponentials (40 in our case) 
# to give us the standard deviation, and then squaring that to get the variance.
sd <- 1/lambda/sqrt(nExp)
theoreticalVariance <- sd^2
```
The theoretical variance is `r theoreticalVariance ` and the sample variance is `r sampleVariance`  
We can see that these values are also very close.

*****************

### Now lets show that the distribution is approximately normal.

As we can see from the plot below, it does indeed show normality.
```{r cache=FALSE, message=FALSE}
qqnorm(exponentialDistributionMeans$means)
qqline(exponentialDistributionMeans$means)
```
